### 2019-09-19 paper/content/abstract.tex
setup basic project structure

% !TeX root = ../paper.tex
\section*{Abstract in deutscher Sprache}

\acresetall{}

\newpage \section*{Abstract in englischer Sprache}

\acresetall{}