### 2019-09-24 paper/content/chapter3/chapter3_1.tex
feat(chore): rough outline

% !TeX root = ../../paper.tex
\subsection{Tools}