### 2019-09-24 paper/content/chapter4.tex

% !TeX root = ../paper.tex
\section{Auswertung} \label{ch:evaluation}