### 2020-01-22 paper/content/chapter2/chapter2_2.tex
Created new file for chapter

% !TeX root = ../../paper.tex
\subsection{Alpha-Beta-Suche}